---
title: "Homework 3"
author: Yuanxin Zhang
output: github_document
---

This is my solution to Homework 3.

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 7,
  fig.asp = .7,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Question 1
### Part 1: Write a short description of the dataset
Note the size and structure of the data, describing some key variables, and giving illustrative examples of observations.
```{r data_q1}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 


### Part 2: Do or answer the following questions
(a) How many aisles are there, and which aisles are the most items ordered from?
```{r aisle_n}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```

There are 134 aisles. Most items are ordered from fresh vegetables aisle.\


(b) Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r aisle_plot}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>%
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>%
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Top three aisles with the most items are fresh vegetables, fresh fruits and packaged vegetables/fruits.\


(c) Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r popular_item}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>%
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```

Top 3 popular items from baking ingredients are Light Brown Sugar, Pure Baking Soda and Cane Sugar; top 3 popular items from dog food care are Snack Sticks Chicken & Rice Recipe Dog Treats, Organix Chicken & Brown Rice Recipe	and Small Dog Biscuits; top 3 popular items from packaged vegetables fruits are Organic Baby Spinach, Organic Raspberries and Organic Blueberries.\


(d) Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r mean_hour}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	) %>% 
	knitr::kable()
```

By looking at table of the mean hour of the day at which products are ordered, Pink Lady Apples are order earlier than Coffee Ice Cream in a day for each day of the day.


# Question 2
### Part 1: Read and clean accelerometers data
Final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes.
```{r readnclean_accel}
accel_df = 
	read_csv(
		"./data/accel_data.csv") %>% 
  pivot_longer(
    activity.1:activity.1440,
    names_to = "minute",
    names_prefix = "activity.",
    values_to = "activity"
  ) %>% 
  mutate(
    weekday = case_when(
      day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday",
      day %in% c("Saturday","Sunday") ~ "weekend",
      TRUE ~ ""
    ),
    minute = as.numeric(minute),
    day = factor(day),
    day = forcats::fct_relevel(day, c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday", "Sunday")),
    weekday = factor(weekday)
  )
```

This dataset contains `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. It is from five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). Variables include week (identifier of week), day_id, day (Monday to Sunday), minute (1440 minutes a day), weekday (weekday vs weekend) and activity, which is the activity count for each minute of a 24-hour day starting at midnight.


### Part 2: Create a table showing the total activity for each day
```{r total_activity}
accel_df %>% 
	group_by(week, day) %>% 
	summarise(total_activity = sum(activity)) %>%
	pivot_wider(
		names_from = day,
		values_from = total_activity
	) %>% 
	knitr::kable()
```

In overall, the participant was more active in weekdays compared to weekends and had two very low total activity counts on the Saturday of Week 4 and Week 5. Total activity counts for Tuesday, Wednesday and Thursday is comparably consistent across weeks.


### Part 3: Make a single-panel plot that shows the 24-hour activity time courses for each day
```{r activity_plot}
accel_df %>% 
  ggplot(aes(x = minute, y = activity, color = day, group = day_id)) + 
  geom_point(alpha = 0.3) + 
  geom_line(alpha = 0.3) + 
  geom_smooth(aes(group = day), se = FALSE) + 
  theme(legend.position = "bottom") +
  labs(
    title = "Activity counts for each minute starting at midnight for each day",
    x = "Minute of a day",
    y = "Activity counts"
  )
```

In overall, the participant is started to get active around 5-6 am on each day, indicating this person wake up around this time. This participant is particularly active around noon on Sunday and around 9pm on Friday.


# Question 3
### Part 1: Read and clean ny_noaa data
```{r read_ny_noaa}
data("ny_noaa")

ny_noaa = ny_noaa %>%
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>%
  mutate(tmax = as.numeric(tmax)/10,
         tmin = as.numeric(tmin)/10,
         prcp = prcp/10,
         ) %>% 
  naniar::replace_with_na(replace = list(snow = -13))

summary(pull(ny_noaa, snow))
```

This dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. It contains records for all New York state weather stations from January 1, 1981 through December 31, 2010. NCEI provides numerous daily variables, including maximum and minimum temperature, total daily precipitation, snowfall, and snow depth. In raw data, there is a record of -13 mm snowfall. Since snowfall cannot be negative, this snowfall record was coded as missing data prior to analysis. There are `r sum(is.na(pull(ny_noaa, prcp)))` missing data in precipitation (mm), `r sum(is.na(pull(ny_noaa, snow)))` missing data in snowfall(mm), `r sum(is.na(pull(ny_noaa, snwd)))` missing data in snow depth(mm), `r sum(is.na(pull(ny_noaa, tmax)))` missing data in maximum temperature (degrees C) and `r sum(is.na(pull(ny_noaa, tmin)))` missing data in minimum temperature (degrees C). Hence, there will be a great loss of data if we omit all record with missing data.

For snowfall, the most commonly observed values are 0 because it only happens at low enough temperature and humidity.


### Part 2: Two-panel plot showing the average max temperature in January and in July in each station across years
```{r max_temp_plot}
max_jan = ny_noaa %>% 
  filter(month == 1) %>% 
  group_by(id, year, month) %>% 
  summarise(tmax_avg = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = tmax_avg, color = id, group = id)) +
  geom_point(alpha = 0.5) +
  geom_path(alpha = 0.3) +
  theme(legend.position = "none") +
  labs(
    title = "Average max temperature in January in each station across years",
    x = "Year",
    y = "Temperature in degree C"
  )

 max_jul = ny_noaa %>% 
  filter(month == 7) %>% 
  group_by(id, year, month) %>% 
  summarise(tmax_avg = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = tmax_avg, color = id, group = id)) +
  geom_point(alpha = 0.5) +
  geom_path(alpha = 0.3) +
  theme(legend.position = "none") +
  labs(
    title = "Average max temperature in January in each station across years",
    x = "Year",
    y = "Temperature in degree C"
  )

 max_jan / max_jul
```

The average max temperature in January usually fluctuate between -10 and 10 degree C. Temperature of January in 1994 and 2004 seem to be lower than other years. Temperature of January across stations vary a lot in 2004.\
The average max temperature in July usually fluctuate between 20 and 30 degree C. Compared to January, the fluctuation of temperature in July across years is smaller. The variation of temperature across stations is also smaller.\
From the two-panel plot, there are a few outliers of average max temperature in both January and July.

### Part 3:
Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
```{r temp_n_snow}
temp = ny_noaa %>%
  drop_na(tmin, tmax) %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) +
  labs(
    title = "Max vs Min temperature",
    x = "Temperature in degree C",
    y = "Density"
  )

snowfall = ny_noaa %>%
  filter(snow > 0 & snow < 100) %>% 
  ggplot(aes(x = snow, y = year, group = as.factor(year))) + 
  geom_density_ridges(alpha = 0.5) +
  labs(
    title = "Distribution of snowfall by year",
    x = "Snowfall (mm)",
    y = "Year"
  )

temp + snowfall
```

The Max vs Min temperature plot indicates that the minimum temperature is ranging from -30 to 30 degree C, with two peaks around 0 and 15 degree C; the maximum temperature is ranging from -15 to 38 degree C, with two peaks around 5 and 25 degree C. There is a big overlap between the distribution of minimum temperature and the distribution of maximum temperature.\
The distribution of snowfall by year plot indicates that there was more snow in earlier years compared to recent years, in overall.


