---
title: "Homework 3"
author: Yuanxin Zhang
output: github_document
---

This is my solution to Homework 3.

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Question 1
### Part 1: Write a short description of the dataset
Note the size and structure of the data, describing some key variables, and giving illustrative examples of observations.
```{r data_q1}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 


### Part 2: Do or answer the following questions
(a) How many aisles are there, and which aisles are the most items ordered from?
```{r aisle_n}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```

There are 134 aisles. Most items are ordered from fresh vegetables aisle.\


(b) Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r aisle_plot}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>%
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>%
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Top three aisles with the most items are fresh vegetables, fresh fruits and packaged vegetables/fruits.\


(c) Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r popular_item}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>%
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```

Top 3 popular items from baking ingredients are Light Brown Sugar, Pure Baking Soda and Cane Sugar; top 3 popular items from dog food care are Snack Sticks Chicken & Rice Recipe Dog Treats, Organix Chicken & Brown Rice Recipe	and Small Dog Biscuits; top 3 popular items from packaged vegetables fruits are Organic Baby Spinach, Organic Raspberries and Organic Blueberries.\


(d) Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r mean_hour}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	) %>% 
	knitr::kable()
```

By looking at table of the mean hour of the day at which products are ordered, Pink Lady Apples are order earlier than Coffee Ice Cream in a day for each day of the day.


# Question 2
### Part 1: Read and clean accelerometers data
```{r readnclean_accel}
accel_df = 
	read_csv(
		"./data/accel_data.csv") %>% 
  pivot_longer(
    activity.1:activity.1440,
    names_to = "minute",
    names_prefix = "activity.",
    values_to = "activity"
  )
```

# Question 3
